---
title: 'Crime in Chicago: A Data Analysis Summary using R'
author: 'Eugeni Vidal'
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    fig_caption: yes
    fig_crop: yes
    keep_tex: yes
    number_sections: yes
    toc_depth: 2
    toc_float: yes
  pdf_document:
    fig_caption: yes
    fig_crop: yes
    fig_height: 3.3
    fig_width: 5.3
    keep_tex: yes
    number_sections: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
header-includes: \usepackage{float} \floatplacement{figure}{H}
---
```{r, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

# Introduction

This report is the first assessment of the **MATH5741M Statistical Theory and Methods** module. Its aims are to summarise a crime dataset from the city of Chicago and answer the following research questions:

- How has crime evolved over time in the city of Chicago? 

- What time of day do most crime occur?

- In which locations of the city is crime more likely to happen?

- Which districts are more potentially dangerous?

# Data and methods

The analysis is done with a sample of the [crime dataset from the Chicago Police Department](https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2) which contains all the crime incidents that occurred in the city of Chicago from 2001 to the present.

We first prepare the data, creating, transforming and cleaning the variables we are interested in. Then, we perform the statistical analysis through line graphs, bar graphs and heat-maps which will answer each of our research questions. 

The report has been done with `Rmarkdown` and it is code reproducible. However, due to space limitation (it should not be more than 6 pages in total) does not include all the code written for its performance. <sup></sup>^[To see the whole code visit https://github.com/eugenividal/Chicago-Crime-Data-Analysis.]

# Results

## Data preparation

First, we activate the libraries we will need to set up the project.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Activate libraries
library(ggplot2)
library(lubridate)
library(zoo)
library(dplyr)
library(knitr)
library(plotly)
```
    
Second, we load the data into the `R` environment.

```{r eval=TRUE, echo=TRUE}
# Read csv in R
dd=read.csv("http://www1.maths.leeds.ac.uk/~charles/math5741/crime.csv",header=T)
```

\pagebreak

Third, we create two new variables `Count` and `Hour` and make some necessary transformations in date variables in order to make `R` understand them. 

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Create a variable count with value 1
dd$Count <- 1
# Convert Date from factor to date
dd$Date <- mdy_hms(dd$Date)
# Extract hour from Date
dd$Hour <- substring(dd$Date, 12,13)
# Drop time from Date
dd$Date <- as.Date(dd$Date, format="%m/%d/%Y")
```

Fourth, we group in larger categories the existent ones in the variables `Primary.Type` and `Location.Description`, and call them `Type_grouped` and `Location_grouped` respectively<sup></sup>^[The code written to group the variables' categories can be seen at https://github.com/eugenividal/Chicago-Crime-Data-Analysis].

```{r, include=FALSE}
# Group Primary.Type categories
dd$Type_grouped[dd$Primary.Type == "THEFT" | dd$Primary.Type == "MOTOR VEHICLE THEFT" ] <- "Theft"

dd$Type_grouped[dd$Primary.Type == "BATTERY"] <- "Batery"

dd$Type_grouped[dd$Primary.Type == "CRIMINAL DAMAGE"] <- "Criminal damage"

dd$Type_grouped[dd$Primary.Type == "NARCOTICS" | dd$Primary.Type == "OTHER NARCOTIC VIOLATION"] <- "Narcotics"

dd$Type_grouped[dd$Primary.Type == "ASSAULT"] <- "Assault"

dd$Type_grouped[dd$Primary.Type == "BURGLARY"] <- "Burglary"

dd$Type_grouped[dd$Primary.Type == "ROBBERY" ] <- "Robery"

dd$Type_grouped[dd$Primary.Type == "ARSON" | dd$Primary.Type == "CONCEALED CARRY LICENSE VIOLATION" | dd$Primary.Type == "CRIMINAL TRESPASS" | dd$Primary.Type == "GAMBLINGS" | dd$Primary.Type == "HUMAN TRAFFICKING" | dd$Primary.Type == "INTERFERENCE WITH PUBLIC OFFICER" | dd$Primary.Type == "INTIMIDATION" | dd$Type == "KIDNAPPING" | dd$Type == "LIQUOR LAW VIOLATION" | dd$Primary.Type == "NON-CRIMINAL" | dd$Primary.Type == "NON - CRIMINAL" | dd$Primary.Type == "OBSCENITY" | dd$Primary.Type == "OFFENSE INVOLVING CHILDREN"| dd$Primary.Type == "PROSTITUTION"| dd$Primary.Type == "PUBLIC INDECENCY"| dd$Primary.Type == "PUBLIC PEACE VIOLATION"| dd$Primary.Type == "STALKING"| dd$Primary.Type == "WEAPONS VIOLATION"| dd$Primary.Type == "HOMICIDE"| dd$Primary.Type == "CRIM SEXUAL ASSAULT" | dd$Primary.Type == "SEX OFFENSE"| dd$Primary.Type == "DECEPTIVE PRACTICE" | dd$Primary.Type == "OTHER OFFENSE"] <- "Others"
```

```{r, include=FALSE}
# Group Location.Description categories
dd$Location_grouped[dd$Location.Description == "\"SCHOOL- PRIVATE- BUILDING\"" | dd$Location.Description ==  "\"SCHOOL- PRIVATE- GROUNDS\"" | dd$Location.Description == "\"SCHOOL- PUBLIC- BUILDING\"" | dd$Location.Description == "\"SCHOOL- PUBLIC- GROUNDS\"" | dd$Location.Description == "COLLEGE/UNIVERSITY GROUNDS" | dd$Location.Description == "COLLEGE/UNIVERSITY RESIDENCE HALL"| dd$Location.Description == "SCHOOL YARD"]<- "School/university"

dd$Location_grouped[dd$Location.Description == "" | dd$Location.Description == "ABANDONED BUILDING"| dd$Location.Description == "AIRCRAFT" | dd$Location.Description == "ANIMAL HOSPITAL" | dd$Location.Description == "ATHLETIC CLUB" | dd$Location.Description == "AUTO" | dd$Location.Description == "BASEMENT"  | dd$Location.Description == "BOAT/WATERCRAFT" | dd$Location.Description == "CHA GROUNDS" | dd$Location.Description == "CHA HALLWAY"  | dd$Location.Description == "CHA HALLWAY/STAIRWELL/ELEVATOR" | dd$Location.Description == "CHURCH" | dd$Location.Description == "CHURCH/SYNAGOGUE/PLACE OF WORSHIP"| dd$Location.Description == "COIN OPERATED MACHINE"| dd$Location.Description == "CONSTRUCTION SITE"| dd$Location.Description == "OTHER"| dd$Location.Description == "CHURCH"| dd$Location.Description == "OTHER RAILROAD PROP / TRAIN DEPOT" | dd$Location.Description =="SEWER" | dd$Location.Description =="STAIRWELL"| dd$Location.Description == "VACANT LOT" | dd$Location.Description =="VACANT LOT/LAND"| dd$Location.Description == "VESTIBULE" | dd$Location.Description =="WOODED AREA" | dd$Location.Description == "CTA STATION" | dd$Location.Description == "CTA BUS STOP"| dd$Location.Description == "CTA TRACKS - RIGHT OF WAY"  | dd$Location.Description == "AIRPORT BUILDING NON-TERMINAL - NON-SECURE AREA" | dd$Location.Description == "AIRPORT BUILDING NON-TERMINAL - SECURE AREA" | dd$Location.Description == "AIRPORT EXTERIOR - NON-SECURE AREA" | dd$Location.Description == "AIRPORT EXTERIOR - SECURE AREA" | dd$Location.Description == "AIRPORT PARKING LOT" | dd$Location.Description ==  "AIRPORT TERMINAL LOWER LEVEL - NON-SECURE AREA" | dd$Location.Description == "AIRPORT TERMINAL LOWER LEVEL - SECURE AREA" | dd$Location.Description == "AIRPORT TERMINAL MEZZANINE - NON-SECURE AREA"| dd$Location.Description == "AIRPORT TERMINAL UPPER LEVEL - NON-SECURE AREA"| dd$Location.Description == "AIRPORT TERMINAL UPPER LEVEL - SECURE AREA"| dd$Location.Description == "AIRPORT TRANSPORTATION SYSTEM (ATS)"| dd$Location.Description == "AIRPORT VENDING ESTABLISHMENT" | dd$Location.Description == "AIRPORT/AIRCRAFT" |dd$Location.Description == "DAY CARE CENTER"| dd$Location.Description == "COMMERCIAL / BUSINESS OFFICE"| dd$Location.Description == "FACTORY" | dd$Location.Description =="FACTORY/MANUFACTURING BUILDING" | dd$Location.Description =="FEDERAL BUILDING"| dd$Location.Description == "FIRE STATION"| dd$Location.Description == "GOVERNMENT BUILDING/PROPERTY"| dd$Location.Description == "HOSPITAL"| dd$Location.Description == "HOSPITAL BUILDING/GROUNDS"| dd$Location.Description == "JAIL / LOCK-UP FACILITY"| dd$Location.Description == "LIBRARY"| dd$Location.Description == "MOVIE HOUSE/THEATER"  | dd$Location.Description =="NURSING HOME/RETIREMENT HOME"| dd$Location.Description == "POOL ROOM" | dd$Location.Description =="SPORTS ARENA/STADIUM"| dd$Location.Description == "WAREHOUSE" | dd$Location.Description == "BANK"| dd$Location.Description == "CREDIT UNION"| dd$Location.Description == "CURRENCY EXCHANGE"| dd$Location.Description == "SAVINGS AND LOAN"]<- "Others"

dd$Location_grouped[dd$Location.Description == "ALLEY" | dd$Location.Description == "BOWLING ALLEY" | dd$Location.Description == "CHA BREEZEWAY" | dd$Location.Description =="HALLWAY"]<- "Alley" 

dd$Location_grouped[dd$Location.Description == "APARTMENT"| dd$Location.Description == "CHA APARTMENT"]<- "Apartment"

dd$Location_grouped[dd$Location.Description == "APPLIANCE STORE" | dd$Location.Description == "BARBERSHOP" | dd$Location.Description == "CAR WASH" | dd$Location.Description == "CLEANING STORE" | dd$Location.Description ==  "CONVENIENCE STORE" | dd$Location.Description =="DEPARTMENT STORE" | dd$Location.Description =="DRUG STORE"| dd$Location.Description == "GARAGE/AUTO REPAIR"| dd$Location.Description == "GAS STATION"| dd$Location.Description == "GAS STATION DRIVE/PROP." | dd$Location.Description =="GROCERY FOOD STORE"| dd$Location.Description == "MEDICAL/DENTAL OFFICE" | dd$Location.Description =="NEWSSTAND"| dd$Location.Description == "OFFICE"| dd$Location.Description == "PAWN SHOP" | dd$Location.Description =="RETAIL STORE"| dd$Location.Description == "SMALL RETAIL STORE"]<- "Store/small business"

dd$Location_grouped[dd$Location.Description == "ATM (AUTOMATIC TELLER MACHINE)" | dd$Location.Description == "BRIDGE"| dd$Location.Description == "DRIVEWAY"| dd$Location.Description == "GANGWAY"| dd$Location.Description == "HIGHWAY/EXPRESSWAY"| dd$Location.Description == "LAKEFRONT/WATERFRONT/RIVERBANK"| dd$Location.Description == "SIDEWALK" | dd$Location.Description == "STREET"]<- "Street"

dd$Location_grouped[dd$Location.Description == "BAR OR TAVERN"| dd$Location.Description == "HOTEL"| dd$Location.Description == "HOTEL/MOTEL"| dd$Location.Description == "RESTAURANT"| dd$Location.Description == "TAVERN"| dd$Location.Description == "TAVERN/LIQUOR STORE"]<- "Restaurant/bar/hotel"

dd$Location_grouped[dd$Location.Description == "CHA PARKING LOT" | dd$Location.Description =="PARK PROPERTY" | dd$Location.Description =="PARKING LOT"| dd$Location.Description == "PARKING LOT/GARAGE(NON.RESID.)" | dd$Location.Description =="POLICE FACILITY/VEH PARKING LOT"]<- "Parking lot"

dd$Location_grouped[dd$Location.Description == "TRAILER" | dd$Location.Description == "TRUCK" | dd$Location.Description == "VEHICLE - DELIVERY TRUCK"| dd$Location.Description ==  "VEHICLE - OTHER RIDE SERVICE"| dd$Location.Description ==  "VEHICLE NON-COMMERCIAL" | dd$Location.Description == "VEHICLE-COMMERCIAL" | dd$Location.Description == "CTA BUS" | dd$Location.Description =="DELIVERY TRUCK" | dd$Location.Description =="TAXICAB" | dd$Location.Description == "CTA TRAIN" | dd$Location.Description == "OTHER COMMERCIAL TRANSPORTATION"]<- "Vehicle" 

dd$Location_grouped[dd$Location.Description == "CTA GARAGE / OTHER PROPERTY" | dd$Location.Description =="DRIVEWAY - RESIDENTIAL" | dd$Location.Description =="GARAGE" | dd$Location.Description =="HOUSE" | dd$Location.Description == "PORCH" | dd$Location.Description =="RESIDENCE" | dd$Location.Description == "RESIDENCE PORCH/HALLWAY"| dd$Location.Description == "RESIDENCE-GARAGE"| dd$Location.Description == "RESIDENTIAL YARD (FRONT/BACK)"| dd$Location.Description == "YARD"]<- "Residence"
```

The next step is to drop all those columns we do not need to answer our research questions.

```{r, echo=TRUE}
# Drop all variables we are not interested in
dd <- dd[, -c(1:2, 4:11, 13:15, 17:18)]
```

Then, we clean the dataset of missing values and remove all values from 2016 - this last year is not complete.

```{r, echo=TRUE}
# Remove NAs
dd <- dd[complete.cases(dd),]
# Remove 2016 rows
dd <- dd[!dd$Year > 2015,]
```

Finally, we show the the dataset ready for exploration.

```{r, echo=TRUE}
# Show first 5 records
head(dd)
```

## Data exploration 

### How has crime evolved over time in the city of Chicago?  

To answer the first question we plot the number of crimes per year from 2001 to 2015 (Figure 1) <sup></sup>^[See visualisations code at https://github.com/eugenividal/Chicago-Crime-Data-Analysis]. The graph shows that crime in the city of Chicago has been decreasing year after year, with a continuous decline. 

```{r fig, fig.cap="Figure 1. Crimes evolution 2001-2015", echo=FALSE, message=FALSE, warning=FALSE, out.extra='',fig.align='center'}
# Create aggregated object
dd_aggr <- aggregate(Count ~ Year, data = dd, FUN = sum)
# Plot the graph 
p1 <- ggplot(dd_aggr, aes(x=Year, y= Count)) + geom_line(colour = "steelblue") + geom_point(colour = "steelblue") + theme_minimal() + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank()) 
p1 <- ggplotly(p1)
p1
```

Figure 2 depicts the annual frequency of crimes per type and their trend. The most common types of crime are Theft and Batery. All types have been falling to a greater or lesser extent. 

```{r fig2, fig.cap="Figure 2. Crimes evolution per type of crime 2001-2015", echo=FALSE, message=FALSE, warning=FALSE, out.extra='',fig.align='center'}
# Create aggregated object
dd_aggr2 <- aggregate(Count ~ Type_grouped + Year, data = dd, FUN = sum)
# Plot the graph
p2 <-ggplot(data=dd_aggr2, aes(x=Year, y=Count, group = Type_grouped, colour = Type_grouped)) +
    geom_line() + geom_point() + theme_minimal() + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())+ theme(legend.title=element_blank())
p2 <- ggplotly(p2)
p2
```

\pagebreak

### What time of day do most crime occur?

The following bar graph (Figure 3) shows the number of crimes increases gradually from 05:00 in the morning (the hour with less crimes) until 20:00 in the evening (the hour with the most crimes). The hours of 12:00 and 00:00 are exceptionally high, at a similar level as 20:00.

```{r, fig3, fig.cap="Figure 3. Crimes per hour", echo=FALSE, out.extra='',fig.align='center'}
# Plot the graph 
p3 <-ggplot(dd, aes(x=Hour))+geom_bar(stat="Count", width=0.8, fill = "steelblue")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Hour", y = "Number of crimes") + theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
p3 <- ggplotly(p3)
p3
```

The heat-map in Figure 4 shows the distribution of number of crimes per hour and type. For example, we can see that the peak hours of Theft and Others are at 00:00, 09:00 and 12:00. Narcotics concentrate between 10:00 to 14:00 and 19:00 to 22:00. Other types are more evenly distributed throughout the day.

```{r fig4, fig.cap="Figure 4. Type of crime vs hour", echo=FALSE, out.extra='', fig.align='center'}
# Create aggregated object
dd_aggr3 <- aggregate(Count ~ Type_grouped + Hour, data = dd, FUN = sum)
# Plot graph
p4 <- ggplot(data = dd_aggr3, aes(x = Hour, y = Type_grouped)) + geom_tile(aes(fill = Count), color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") 
p4<- p4 + theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank()) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 6),
        axis.text.y = element_text(size= 8),
        axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) 
p4 <- ggplotly(p4)
p4
```

\pagebreak

### In which locations of the city is crime more likely to happen? 

As is illustrated by Figure 5, most crimes happen in the street, followed by Residences and Apartments.

```{r fig5, fig.cap="Figure 5. Crimes per location", echo=FALSE, out.extra='',fig.align='center'}
# Create aggregated object
dd_aggr4 <- aggregate(Count ~ Location_grouped, data = dd, FUN = sum)
# Order values
dd_aggr4$Location_grouped <- factor(dd_aggr4$Location_grouped, levels = dd_aggr4$Location_grouped[order(-dd_aggr4$Count)])
# Plot the graph 
p5 <-ggplot(dd_aggr4, aes(x = Location_grouped, y = Count)) + theme_minimal() + geom_bar(stat="identity", width=0.7, fill = "steelblue") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(x = "Location", y = "Number of crimes")+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
p5 <- ggplotly(p5)
p5
```

If we visualise the distribution of crimes per location and type (Figure 6), we can see that some types occur in specific locations. For instance, Robery is recorded almost enterily in the Street, as well as Narcotics. However, Burglary and Others are registered particularly in Reseidences or Apartments. What makes sense.

```{r fig6, fig.cap="Figure 6. Type of crime vs location", echo=FALSE, out.extra='',fig.align='center'}
# Create aggregated object
dd_aggr5 <- aggregate(Count ~  Type_grouped + Location_grouped, data = dd, FUN = sum)
# Plot the graph
p6 <- ggplot(data = dd_aggr5, aes(x = Location_grouped, y = Type_grouped)) + geom_tile(aes(fill = Count), color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") 
p6 <-p6+ theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank()) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 6),
        axis.text.y = element_text(size= 8),
        axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) 
p6 <- ggplotly(p6)
p6
```

\pagebreak

### Which districts are more potentially dangerous?

Finally, in Figure 7, we visualise the number of crimes per districts. The most dangerous district seems number 8, with more than 30,000 records in the 15 years, while district 20 with less than 10,000 seems the safest. 

```{r fig7, echo=FALSE, fig.cap="Figure 7. Crimes per district",fig.align='center'}
# Remove values districts 21 and 31
dd$District<- as.factor(dd$District)
dd_sub <- subset(dd, District!="21" & District!="31")
# Create aggregated object
dd_aggr6 <- aggregate(Count ~ District, data = dd_sub, FUN = sum)
# Order values
dd_aggr6$District <- factor(dd_aggr6$District, levels = dd_aggr6$District[order(-dd_aggr6$Count)])
# Plot the graph 
p7<-ggplot(dd_aggr6, aes(x=District, y = Count)) + theme_minimal() + geom_bar(stat="identity", width=0.7, fill = "steelblue") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(x = "District", y = "Number of crimes") + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
p7 <- ggplotly(p7)
p7
```

Figure 8 ilustrates some interesting findings in the relations between the type of crime and districts where they occurred. For example, we can see that districts 1, 8, 12 and 18 are particularly dangerous in terms of Theft, that districts 7 and 11 stand out in terms of Batery, but above all that Narcotics crime concentrates in district 11 and 15.

```{r fig8, fig.cap="Figure 8. Type of crime vs district", echo=FALSE, out.extra='',fig.align='center'}
# Create aggregated object
dd_aggr7 <- aggregate(Count ~ Type_grouped + District, data = dd_sub, FUN = sum)
# Plot the graph
p8<-ggplot(data = dd_aggr7, aes(x = District, y = Type_grouped)) +
  geom_tile(aes(fill = Count), color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue")  
p8<- p8+ theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank()) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 6),
        axis.text.y = element_text(size= 8),
        axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) 
p8 <- ggplotly(p8)
p8
```
