---
title: "Chicago Crime: Data Analysis and \n Visualisations using R"
author: 'Student ID: 201081646'
date: "November 3, 2018"
output:
  pdf_document:
    fig_caption: yes
    fig_crop: yes
    fig_height: 3.5
    fig_width: 5.5
    keep_tex: yes
    number_sections: yes
    toc_depth: 2
  html_document:
    fig_caption: yes
    theme: journal
    toc: yes
    toc_depth: 2
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
---
```{r, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

# Introduction

This report is the first assessment for the **MATH5741M Statistical Theory and Methods** module. Its objective is to summarise statistically a dataset sample of crimes in the city of Chicago and due to limit has restic to answer the following research questions:

- How has crime evolved over time in the city of Chicago? 

- When time of day do most crime occur?

- In which locations is crime more likely to happen?

# Data and methods

The dataset analysed is a sample of the [original data of crimes extracted from the Chicago Police Department](https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2) which content the crimes that occurred in the city of Chicago from 2001 to present.

For the analysis, first, we prepare the data creating, transforming and simplifying variables, as well as cleaning the dataset keeping the variables we are interested in. Secondly, to answer our research questions we perform the an analysis based on line graphs and heat-maps. Finally, we summarise the findings.

The report has been done with `Rmarkdown` but unfortunately does not include all the `R` code cells written for its performance<sup></sup>^[In this report it is not included the code used to group categories in variables `Primary.Type` and `Location.Description`, neither the code used to generate the visualisations]. However, it is available for consultation in this link https://github.com/eugenividal/Chicago-Crime-Data-Analysis.

# Results

## Data preparation

First, we load the libraries we will need for the project and get the data into the `R` environment.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# load libraries
library(ggplot2)
library(ggmap)
library(lubridate)
library(scales)
library(zoo)
library(dplyr)
library(knitr)
```

```{r eval=TRUE, echo=TRUE}
# Read csv in R
dd=read.csv("http://www1.maths.leeds.ac.uk/~charles/math5741/crime.csv",header=T)
```

Second, we create the new variables (`Count`, `Month_year`, `Hour`) based on the existing ones, and give them the right format for later exploration. 

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Create a variable count with value 1
dd$Count <- 1
# Convert Date from factor to date
dd$Date <- mdy_hms(dd$Date)
# Extract hour from Date
dd$Hour <- substring(dd$Date, 12,13)
# Drop time from Date
dd$Date <- as.Date(dd$Date, format="%m/%d/%Y")
# Drop days from Date
dd$Month_year <- as.Date(as.yearmon(dd$Date, "%Y-%m"))
```

Third, we group the categories of the variables `Primary.Type` and `Location.Description` to simplify the analysis and call them `Type_grouped` and `Location_grouped` respectively.

```{r, include=FALSE}
# Group Primary.Type categories
dd$Type_grouped[dd$Primary.Type == "THEFT" | dd$Primary.Type == "MOTOR VEHICLE THEFT" ] <- "Theft"

dd$Type_grouped[dd$Primary.Type == "BATTERY"] <- "Batery"

dd$Type_grouped[dd$Primary.Type == "CRIMINAL DAMAGE"] <- "Criminal damage"

dd$Type_grouped[dd$Primary.Type == "NARCOTICS" | dd$Primary.Type == "OTHER NARCOTIC VIOLATION"] <- "Narcotics"

dd$Type_grouped[dd$Primary.Type == "OTHER OFFENSE"] <- "Other offense"

dd$Type_grouped[dd$Primary.Type == "ASSAULT"] <- "Assault"

dd$Type_grouped[dd$Primary.Type == "BURGLARY"] <- "Burglary"

dd$Type_grouped[dd$Primary.Type == "ROBBERY" ] <- "Robery"

dd$Type_grouped[dd$Primary.Type == "DECEPTIVE PRACTICE"] <- "Deceptive practice"

dd$Type_grouped[dd$Primary.Type == "ARSON" | dd$Primary.Type == "CONCEALED CARRY LICENSE VIOLATION" | dd$Primary.Type == "CRIMINAL TRESPASS" | dd$Primary.Type == "GAMBLINGS" | dd$Primary.Type == "HUMAN TRAFFICKING" | dd$Primary.Type == "INTERFERENCE WITH PUBLIC OFFICER" | dd$Primary.Type == "INTIMIDATION" | dd$Type == "KIDNAPPING" | dd$Type == "LIQUOR LAW VIOLATION" | dd$Primary.Type == "NON-CRIMINAL" | dd$Primary.Type == "NON - CRIMINAL" | dd$Primary.Type == "OBSCENITY" | dd$Primary.Type == "OFFENSE INVOLVING CHILDREN"| dd$Primary.Type == "PROSTITUTION"| dd$Primary.Type == "PUBLIC INDECENCY"| dd$Primary.Type == "PUBLIC PEACE VIOLATION"| dd$Primary.Type == "STALKING"| dd$Primary.Type == "WEAPONS VIOLATION"| dd$Primary.Type == "HOMICIDE"| dd$Primary.Type == "CRIM SEXUAL ASSAULT" | dd$Primary.Type == "SEX OFFENSE"] <- "Others"
```

```{r, include=FALSE}
# Group Location.Description categories
dd$Location_grouped[dd$Location.Description == "\"SCHOOL- PRIVATE- BUILDING\"" | dd$Location.Description ==  "\"SCHOOL- PRIVATE- GROUNDS\"" | dd$Location.Description == "\"SCHOOL- PUBLIC- BUILDING\"" | dd$Location.Description == "\"SCHOOL- PUBLIC- GROUNDS\"" | dd$Location.Description == "COLLEGE/UNIVERSITY GROUNDS" | dd$Location.Description == "COLLEGE/UNIVERSITY RESIDENCE HALL"| dd$Location.Description == "SCHOOL YARD"]<- "School/university"

dd$Location_grouped[dd$Location.Description == "" | dd$Location.Description == "ABANDONED BUILDING"| dd$Location.Description == "AIRCRAFT" | dd$Location.Description == "ANIMAL HOSPITAL" | dd$Location.Description == "ATHLETIC CLUB" | dd$Location.Description == "AUTO" | dd$Location.Description == "BASEMENT"  | dd$Location.Description == "BOAT/WATERCRAFT" | dd$Location.Description == "CHA GROUNDS" | dd$Location.Description == "CHA HALLWAY"  | dd$Location.Description == "CHA HALLWAY/STAIRWELL/ELEVATOR" | dd$Location.Description == "CHURCH" | dd$Location.Description == "CHURCH/SYNAGOGUE/PLACE OF WORSHIP"| dd$Location.Description == "COIN OPERATED MACHINE"| dd$Location.Description == "CONSTRUCTION SITE"| dd$Location.Description == "OTHER"| dd$Location.Description == "CHURCH"| dd$Location.Description == "OTHER RAILROAD PROP / TRAIN DEPOT" | dd$Location.Description =="SEWER" | dd$Location.Description =="STAIRWELL"| dd$Location.Description == "VACANT LOT" | dd$Location.Description =="VACANT LOT/LAND"| dd$Location.Description == "VESTIBULE" | dd$Location.Description =="WOODED AREA" | dd$Location.Description == "CTA STATION" | dd$Location.Description == "CTA BUS STOP"| dd$Location.Description == "CTA TRACKS - RIGHT OF WAY"  | dd$Location.Description == "AIRPORT BUILDING NON-TERMINAL - NON-SECURE AREA" | dd$Location.Description == "AIRPORT BUILDING NON-TERMINAL - SECURE AREA" | dd$Location.Description == "AIRPORT EXTERIOR - NON-SECURE AREA" | dd$Location.Description == "AIRPORT EXTERIOR - SECURE AREA" | dd$Location.Description == "AIRPORT PARKING LOT" | dd$Location.Description ==  "AIRPORT TERMINAL LOWER LEVEL - NON-SECURE AREA" | dd$Location.Description == "AIRPORT TERMINAL LOWER LEVEL - SECURE AREA" | dd$Location.Description == "AIRPORT TERMINAL MEZZANINE - NON-SECURE AREA"| dd$Location.Description == "AIRPORT TERMINAL UPPER LEVEL - NON-SECURE AREA"| dd$Location.Description == "AIRPORT TERMINAL UPPER LEVEL - SECURE AREA"| dd$Location.Description == "AIRPORT TRANSPORTATION SYSTEM (ATS)"| dd$Location.Description == "AIRPORT VENDING ESTABLISHMENT" | dd$Location.Description == "AIRPORT/AIRCRAFT" |dd$Location.Description == "DAY CARE CENTER"| dd$Location.Description == "COMMERCIAL / BUSINESS OFFICE"| dd$Location.Description == "FACTORY" | dd$Location.Description =="FACTORY/MANUFACTURING BUILDING" | dd$Location.Description =="FEDERAL BUILDING"| dd$Location.Description == "FIRE STATION"| dd$Location.Description == "GOVERNMENT BUILDING/PROPERTY"| dd$Location.Description == "HOSPITAL"| dd$Location.Description == "HOSPITAL BUILDING/GROUNDS"| dd$Location.Description == "JAIL / LOCK-UP FACILITY"| dd$Location.Description == "LIBRARY"| dd$Location.Description == "MOVIE HOUSE/THEATER"  | dd$Location.Description =="NURSING HOME/RETIREMENT HOME"| dd$Location.Description == "POOL ROOM" | dd$Location.Description =="SPORTS ARENA/STADIUM"| dd$Location.Description == "WAREHOUSE" | dd$Location.Description == "BANK"| dd$Location.Description == "CREDIT UNION"| dd$Location.Description == "CURRENCY EXCHANGE"| dd$Location.Description == "SAVINGS AND LOAN"]<- "Others"

dd$Location_grouped[dd$Location.Description == "ALLEY" | dd$Location.Description == "BOWLING ALLEY" | dd$Location.Description == "CHA BREEZEWAY" | dd$Location.Description =="HALLWAY"]<- "Alley" 

dd$Location_grouped[dd$Location.Description == "APARTMENT"| dd$Location.Description == "CHA APARTMENT"]<- "Apartment"

dd$Location_grouped[dd$Location.Description == "APPLIANCE STORE" | dd$Location.Description == "BARBERSHOP" | dd$Location.Description == "CAR WASH" | dd$Location.Description == "CLEANING STORE" | dd$Location.Description ==  "CONVENIENCE STORE" | dd$Location.Description =="DEPARTMENT STORE" | dd$Location.Description =="DRUG STORE"| dd$Location.Description == "GARAGE/AUTO REPAIR"| dd$Location.Description == "GAS STATION"| dd$Location.Description == "GAS STATION DRIVE/PROP." | dd$Location.Description =="GROCERY FOOD STORE"| dd$Location.Description == "MEDICAL/DENTAL OFFICE" | dd$Location.Description =="NEWSSTAND"| dd$Location.Description == "OFFICE"| dd$Location.Description == "PAWN SHOP" | dd$Location.Description =="RETAIL STORE"| dd$Location.Description == "SMALL RETAIL STORE"]<- "Store/small business"

dd$Location_grouped[dd$Location.Description == "ATM (AUTOMATIC TELLER MACHINE)" | dd$Location.Description == "BRIDGE"| dd$Location.Description == "DRIVEWAY"| dd$Location.Description == "GANGWAY"| dd$Location.Description == "HIGHWAY/EXPRESSWAY"| dd$Location.Description == "LAKEFRONT/WATERFRONT/RIVERBANK"| dd$Location.Description == "SIDEWALK" | dd$Location.Description == "STREET"]<- "Street"

dd$Location_grouped[dd$Location.Description == "BAR OR TAVERN"| dd$Location.Description == "HOTEL"| dd$Location.Description == "HOTEL/MOTEL"| dd$Location.Description == "RESTAURANT"| dd$Location.Description == "TAVERN"| dd$Location.Description == "TAVERN/LIQUOR STORE"]<- "Restaurant/bar/hotel"

dd$Location_grouped[dd$Location.Description == "CHA PARKING LOT" | dd$Location.Description =="PARK PROPERTY" | dd$Location.Description =="PARKING LOT"| dd$Location.Description == "PARKING LOT/GARAGE(NON.RESID.)" | dd$Location.Description =="POLICE FACILITY/VEH PARKING LOT"]<- "Parking lot"

dd$Location_grouped[dd$Location.Description == "TRAILER" | dd$Location.Description == "TRUCK" | dd$Location.Description == "VEHICLE - DELIVERY TRUCK"| dd$Location.Description ==  "VEHICLE - OTHER RIDE SERVICE"| dd$Location.Description ==  "VEHICLE NON-COMMERCIAL" | dd$Location.Description == "VEHICLE-COMMERCIAL" | dd$Location.Description == "CTA BUS" | dd$Location.Description =="DELIVERY TRUCK" | dd$Location.Description =="TAXICAB" | dd$Location.Description == "CTA TRAIN" | dd$Location.Description == "OTHER COMMERCIAL TRANSPORTATION"]<- "Vehicle" 

dd$Location_grouped[dd$Location.Description == "CTA GARAGE / OTHER PROPERTY" | dd$Location.Description =="DRIVEWAY - RESIDENTIAL" | dd$Location.Description =="GARAGE" | dd$Location.Description =="HOUSE" | dd$Location.Description == "PORCH" | dd$Location.Description =="RESIDENCE" | dd$Location.Description == "RESIDENCE PORCH/HALLWAY"| dd$Location.Description == "RESIDENCE-GARAGE"| dd$Location.Description == "RESIDENTIAL YARD (FRONT/BACK)"| dd$Location.Description == "YARD"]<- "Residence"
```

Next, we drop all variables we do not need to answer our research questions.

```{r, echo=TRUE}
# Drop all variables we are not interested in
dd <- dd[, -c(1:2, 4:11, 13:18)]
```

Then, we clean the dataset of missing values.

```{r, echo=TRUE}
# Remove NAs
dd <- dd[complete.cases(dd),]
```

Finally, we show the the dataset ready for exploration.

```{r, echo=TRUE}
# Show first 5 records
head(dd)
```

## Data exploration

### How has crime evolved over time in the city of Chicago?  

To answer the first question we plot in Figure 1 the number of crimes per month from 2001 to the present. The graph shows that crime in the city of Chicago has been decreasing consistently over the whole period. The wave-shape of the graph also shows that there is a clear periodic pattern per months of the year.

```{r fig, fig.cap="Crimes evolution", echo=FALSE, message=FALSE, warning=FALSE}
# Plot the graph 
ggplot(dd, aes(x=Month_year))+geom_line(color='steelblue', stat="Count") + theme_minimal() + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank()) 
```

In Figure 2 we plot the same graph per type of crime ordered per importance, being Theft the most significant and deceptive practice the less. Except the deceptive practice crime, which at present keeps in similar values as in 2001, the rest of type have been falling to a greater or lesser extent. 

```{r, include=FALSE}
# create the function multiplot
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)
  # make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }
 if (numPlots==1) {
    print(plots[[1]])
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

```{r}
ggplot(data=dd, aes(x=Month_year, y=Count, group = Type_grouped, colour = Type_grouped)) +
    geom_line() +
    geom_point( size=4, shape=21, fill="white")
```


```{r, fig2, fig.cap="Evolution per type of crime", echo=FALSE, message=FALSE, warning=FALSE, out.extra='',fig.align='center'}
# Create a graph with each type evolution
p1 <- ggplot(subset(dd, Type_grouped == "Theft"), aes(x=Month_year))+geom_line(color='steelblue', stat="count")+
labs(y = "Theft") + theme_minimal() + theme(axis.title.x=element_blank()) + theme(text = element_text(size=8))

p2 <- ggplot(subset(dd, Type_grouped == "Batery"), aes(x=Month_year))+geom_line(color='steelblue',stat="count")+
labs(y = "Batery") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p3 <- ggplot(subset(dd, Type_grouped == "Criminal damage"), aes(x=Month_year))+geom_line(color='steelblue', stat="count")+
labs(y = "Criminal\n damage") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p4 <- ggplot(subset(dd, Type_grouped == "Narcotics"), aes(x=Month_year))+geom_line(color='steelblue', stat="count")+
labs(y = "Narcotics") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p5 <- ggplot(subset(dd, Type_grouped == "Others"), aes(x=Month_year))+geom_line(color='steelblue', stat="count")+
labs(y = "Others") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p6 <- ggplot(subset(dd, Type_grouped == "Other offense"), aes(x=Month_year))+geom_line(color='steelblue', stat="count")+
labs(y = "Other\n offense") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p7 <- ggplot(subset(dd, Type_grouped == "Assault"),  aes(x=Month_year))+geom_line(color='steelblue', stat="count") +
labs(y = "Assault") + theme_minimal() + theme(axis.title.x=element_blank()) + theme(text = element_text(size=8))

p8 <- ggplot(subset(dd, Type_grouped == "Burglary"), aes(x=Month_year))+geom_line(color='steelblue', stat="count")+
labs(y = "Burglary") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p9 <- ggplot(subset(dd, Type_grouped == "Robery"), aes(x=Month_year))+geom_line(color='steelblue', stat="count")+
labs(y = "Robery") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p10 <- ggplot(subset(dd, Type_grouped == "Deceptive practice"), aes(x=Month_year))+geom_line(color='steelblue', stat="count")+
labs(y = "Deceptive\n practice") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, cols=2) 
```

### What time of day do most crime occur?

The crimes are concentrated in hours

```{r, fig3, fig.cap="Crimes per hour", echo=FALSE, out.extra='',fig.align='center'}
# Plot the graph 
ggplot(dd, aes(x=Hour))+geom_bar(stat="Count", width=0.8, fill = "steelblue")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Hour", y = "Number of crimes") + theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
```

Figure 3 a head map between type of crime and hour. The peak hour of crimes morning, other crimes peak during the and the final group of crimes peak at night.

```{r fig9, fig.cap="Type of crime vs hour", echo=FALSE, out.extra='', fig.align='center'}
# add variable: performance
count_h <- aggregate(Count ~ Type_grouped + Hour, data = dd, FUN = sum)
# Plot graph
p1 <- ggplot(data = count_h, aes(x = Hour, y = Type_grouped)) + geom_tile(aes(fill = Count), color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") 
p1 + theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank()) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 6),
        axis.text.y = element_text(size= 8),
        axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) 
```

### In which locations is crime more likely to happen? 

These crimes are concentrated in Streets, give percentage.

```{r fig7, fig.cap="Crimes per location", echo=FALSE, out.extra='',fig.align='center'}
# create aggregated object
dd_agg_l <- aggregate(Count ~ Location_grouped, data = dd, FUN = sum)
# order values
dd_agg_l$Location_grouped <- factor(dd_agg_l$Location_grouped, levels = dd_agg_l$Location_grouped[order(-dd_agg_l$Count)])
# plot the graph 
ggplot(dd_agg_l, aes(x = Location_grouped, y = Count)) + theme_minimal() + geom_bar(stat="identity", width=0.7, fill = "steelblue") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(x = "Location", y = "Number of crimes")+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
```

Figure 4 shows that some locations are more likely per type of crime. For theft for instance it is better this or that.

```{r fig11, fig.cap="Type of crime vs location", echo=FALSE, out.extra='',fig.align='center'}
# add variable: performance
count_l <- aggregate(Count ~  Type_grouped + Location_grouped, data = dd, FUN = sum)

p2 <- ggplot(data = count_l, aes(x = Location_grouped, y = Type_grouped)) + geom_tile(aes(fill = Count), color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") 
p2+ theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank()) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 6),
        axis.text.y = element_text(size= 8),
        axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) 
```

# Conclusions


