---
title: 'Chicago Crime Data Analysis'
author: '201081646'
date: "November 3, 2018"
output:
  pdf_document: 
    fig_caption: yes
    fig_height: 2.5
    fig_width: 5
    keep_tex: yes
    number_sections: yes
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# load libraries
library(ggplot2)
library(ggmap)
library(lubridate)
library(scales)
library(zoo)
library(dplyr)
```

# Introduction

This is the first assessment for the **MATH5741M Statistical Theory and Methods** module. Its objective is to summarise statistically a crime dataset sample from the city of Chicago and highlight the key findings.

# Data and methods

The dataset analysed is a 500,000 rows sample of the original data of crimes that occurred in the City of Chicago from 2001 to present extracted from the Chicago Police Department.

The report is structured as follows. First, we prepare the data  creating, transforming and cleaning the variables we are interested in. Secondly, we explore them in an analysis that we divide between univariable based on line and bar graphs and a multianalysys based on heatmaps. Finally, in counclusions we sumarise the main findings.

The document describes not only the statistical process followed but also the most important code used to carry it out. Fro consulting the whole can be consulted in this link.

# Results

## Data preparation

First, we load the data into the `R` environment.

```{r eval=TRUE, echo=TRUE}
# read csv in R
dd=read.csv("http://www1.maths.leeds.ac.uk/~charles/math5741/crime.csv",header=T)
```

Second, we create new variables (`count`, `hour`,  `Month_Yr`, `Month`, and `weekday`) based on the existing ones, and give them the right format for later explotation. 

```{r, include=FALSE}
# Create a variable count with value 1
dd$count <- 1

# Convert Date from factor to date
dd$Date <- mdy_hms(dd$Date)

# Extract hour from Date
dd$hour <- substring(dd$Date, 12,13)

# Drop time from Date
dd$Date <- as.Date(dd$Date, format="%m/%d/%Y")

# Create a month year
dd$Month_Yr <- format(as.Date(dd$Date), "%Y-%m")

# Extract month  starting on January)
dd$mon <- month(dd$Date, label=TRUE)

# Extract weekdays 
dd$weekday <- wday(dd$Date, label=TRUE)
```

```{r, include=FALSE}
# change format
dd$hour <- as.factor(dd$hour)
dd$Month_Yr <- as.Date(as.yearmon(dd$Month_Yr, "%Y-%m"))
dd$mon <- as.factor(dd$mon)
dd$weekday <- as.factor(dd$weekday)
dd$Primary.Type <- as.factor(dd$Primary.Type)
dd$Location.Description <- as.factor(dd$Location.Description)
dd$District<- as.factor(dd$District)
```

Third, we simplify the variables `Primary.Type` and `Location.Description` grouping their categories and call them `Type_grouped` and `Location_grouped` respectivelly.

```{r, include=FALSE}
# Group Primary.Type categories
dd$Type_grouped[dd$Primary.Type == "THEFT" | dd$Primary.Type == "MOTOR VEHICLE THEFT" ] <- "Theft"

dd$Type_grouped[dd$Primary.Type == "BATTERY"] <- "Batery"

dd$Type_grouped[dd$Primary.Type == "CRIMINAL DAMAGE"] <- "Criminal damage"

dd$Type_grouped[dd$Primary.Type == "NARCOTICS" | dd$Primary.Type == "OTHER NARCOTIC VIOLATION"] <- "Narcotics"

dd$Type_grouped[dd$Primary.Type == "OTHER OFFENSE"] <- "Other offense"

dd$Type_grouped[dd$Primary.Type == "ASSAULT"] <- "Assault"

dd$Type_grouped[dd$Primary.Type == "BURGLARY"] <- "Burglary"

dd$Type_grouped[dd$Primary.Type == "ROBBERY" ] <- "Robery"

dd$Type_grouped[dd$Primary.Type == "DECEPTIVE PRACTICE"] <- "Deceptive practice"

dd$Type_grouped[dd$Primary.Type == "ARSON" | dd$Primary.Type == "CONCEALED CARRY LICENSE VIOLATION" | dd$Primary.Type == "CRIMINAL TRESPASS" | dd$Primary.Type == "GAMBLINGS" | dd$Primary.Type == "HUMAN TRAFFICKING" | dd$Primary.Type == "INTERFERENCE WITH PUBLIC OFFICER" | dd$Primary.Type == "INTIMIDATION" | dd$Type == "KIDNAPPING" | dd$Type == "LIQUOR LAW VIOLATION" | dd$Primary.Type == "NON-CRIMINAL" | dd$Primary.Type == "NON - CRIMINAL" | dd$Primary.Type == "OBSCENITY" | dd$Primary.Type == "OFFENSE INVOLVING CHILDREN"| dd$Primary.Type == "PROSTITUTION"| dd$Primary.Type == "PUBLIC INDECENCY"| dd$Primary.Type == "PUBLIC PEACE VIOLATION"| dd$Primary.Type == "STALKING"| dd$Primary.Type == "WEAPONS VIOLATION"| dd$Primary.Type == "HOMICIDE"| dd$Primary.Type == "CRIM SEXUAL ASSAULT" | dd$Primary.Type == "SEX OFFENSE"] <- "Others"
```

```{r, include=FALSE}
# Group Location.Description categories
dd$Location_grouped[dd$Location.Description == "\"SCHOOL- PRIVATE- BUILDING\"" | dd$Location.Description ==  "\"SCHOOL- PRIVATE- GROUNDS\"" | dd$Location.Description == "\"SCHOOL- PUBLIC- BUILDING\"" | dd$Location.Description == "\"SCHOOL- PUBLIC- GROUNDS\"" | dd$Location.Description == "COLLEGE/UNIVERSITY GROUNDS" | dd$Location.Description == "COLLEGE/UNIVERSITY RESIDENCE HALL"| dd$Location.Description == "SCHOOL YARD"]<- "School or university"

dd$Location_grouped[dd$Location.Description == "" | dd$Location.Description == "ABANDONED BUILDING"| dd$Location.Description == "AIRCRAFT" | dd$Location.Description == "ANIMAL HOSPITAL" | dd$Location.Description == "ATHLETIC CLUB" | dd$Location.Description == "AUTO" | dd$Location.Description == "BASEMENT"  | dd$Location.Description == "BOAT/WATERCRAFT" | dd$Location.Description == "CHA GROUNDS" | dd$Location.Description == "CHA HALLWAY"  | dd$Location.Description == "CHA HALLWAY/STAIRWELL/ELEVATOR" | dd$Location.Description == "CHURCH" | dd$Location.Description == "CHURCH/SYNAGOGUE/PLACE OF WORSHIP"| dd$Location.Description == "COIN OPERATED MACHINE"| dd$Location.Description == "CONSTRUCTION SITE"| dd$Location.Description == "OTHER"| dd$Location.Description == "CHURCH"| dd$Location.Description == "OTHER RAILROAD PROP / TRAIN DEPOT" | dd$Location.Description =="SEWER" | dd$Location.Description =="STAIRWELL"| dd$Location.Description == "VACANT LOT" | dd$Location.Description =="VACANT LOT/LAND"| dd$Location.Description == "VESTIBULE" | dd$Location.Description =="WOODED AREA" | dd$Location.Description == "CTA STATION" | dd$Location.Description == "CTA BUS STOP"| dd$Location.Description == "CTA TRACKS - RIGHT OF WAY"  | dd$Location.Description == "AIRPORT BUILDING NON-TERMINAL - NON-SECURE AREA" | dd$Location.Description == "AIRPORT BUILDING NON-TERMINAL - SECURE AREA" | dd$Location.Description == "AIRPORT EXTERIOR - NON-SECURE AREA" | dd$Location.Description == "AIRPORT EXTERIOR - SECURE AREA" | dd$Location.Description == "AIRPORT PARKING LOT" | dd$Location.Description ==  "AIRPORT TERMINAL LOWER LEVEL - NON-SECURE AREA" | dd$Location.Description == "AIRPORT TERMINAL LOWER LEVEL - SECURE AREA" | dd$Location.Description == "AIRPORT TERMINAL MEZZANINE - NON-SECURE AREA"| dd$Location.Description == "AIRPORT TERMINAL UPPER LEVEL - NON-SECURE AREA"| dd$Location.Description == "AIRPORT TERMINAL UPPER LEVEL - SECURE AREA"| dd$Location.Description == "AIRPORT TRANSPORTATION SYSTEM (ATS)"| dd$Location.Description == "AIRPORT VENDING ESTABLISHMENT" | dd$Location.Description == "AIRPORT/AIRCRAFT" |dd$Location.Description == "DAY CARE CENTER"| dd$Location.Description == "COMMERCIAL / BUSINESS OFFICE"| dd$Location.Description == "FACTORY" | dd$Location.Description =="FACTORY/MANUFACTURING BUILDING" | dd$Location.Description =="FEDERAL BUILDING"| dd$Location.Description == "FIRE STATION"| dd$Location.Description == "GOVERNMENT BUILDING/PROPERTY"| dd$Location.Description == "HOSPITAL"| dd$Location.Description == "HOSPITAL BUILDING/GROUNDS"| dd$Location.Description == "JAIL / LOCK-UP FACILITY"| dd$Location.Description == "LIBRARY"| dd$Location.Description == "MOVIE HOUSE/THEATER"  | dd$Location.Description =="NURSING HOME/RETIREMENT HOME"| dd$Location.Description == "POOL ROOM" | dd$Location.Description =="SPORTS ARENA/STADIUM"| dd$Location.Description == "WAREHOUSE" | dd$Location.Description == "BANK"| dd$Location.Description == "CREDIT UNION"| dd$Location.Description == "CURRENCY EXCHANGE"| dd$Location.Description == "SAVINGS AND LOAN"]<- "Others"

dd$Location_grouped[dd$Location.Description == "ALLEY" | dd$Location.Description == "BOWLING ALLEY" | dd$Location.Description == "CHA BREEZEWAY" | dd$Location.Description =="HALLWAY"]<- "Alley" 

dd$Location_grouped[dd$Location.Description == "APARTMENT"| dd$Location.Description == "CHA APARTMENT"]<- "Apartment"

dd$Location_grouped[dd$Location.Description == "APPLIANCE STORE" | dd$Location.Description == "BARBERSHOP" | dd$Location.Description == "CAR WASH" | dd$Location.Description == "CLEANING STORE" | dd$Location.Description ==  "CONVENIENCE STORE" | dd$Location.Description =="DEPARTMENT STORE" | dd$Location.Description =="DRUG STORE"| dd$Location.Description == "GARAGE/AUTO REPAIR"| dd$Location.Description == "GAS STATION"| dd$Location.Description == "GAS STATION DRIVE/PROP." | dd$Location.Description =="GROCERY FOOD STORE"| dd$Location.Description == "MEDICAL/DENTAL OFFICE" | dd$Location.Description =="NEWSSTAND"| dd$Location.Description == "OFFICE"| dd$Location.Description == "PAWN SHOP" | dd$Location.Description =="RETAIL STORE"| dd$Location.Description == "SMALL RETAIL STORE"]<- "Store/small business"

dd$Location_grouped[dd$Location.Description == "ATM (AUTOMATIC TELLER MACHINE)" | dd$Location.Description == "BRIDGE"| dd$Location.Description == "DRIVEWAY"| dd$Location.Description == "GANGWAY"| dd$Location.Description == "HIGHWAY/EXPRESSWAY"| dd$Location.Description == "LAKEFRONT/WATERFRONT/RIVERBANK"| dd$Location.Description == "SIDEWALK" | dd$Location.Description == "STREET"]<- "Street"

dd$Location_grouped[dd$Location.Description == "BAR OR TAVERN"| dd$Location.Description == "HOTEL"| dd$Location.Description == "HOTEL/MOTEL"| dd$Location.Description == "RESTAURANT"| dd$Location.Description == "TAVERN"| dd$Location.Description == "TAVERN/LIQUOR STORE"]<- "Restaurant/bar/hotel"

dd$Location_grouped[dd$Location.Description == "CHA PARKING LOT" | dd$Location.Description =="PARK PROPERTY" | dd$Location.Description =="PARKING LOT"| dd$Location.Description == "PARKING LOT/GARAGE(NON.RESID.)" | dd$Location.Description =="POLICE FACILITY/VEH PARKING LOT"]<- "Parking lot"

dd$Location_grouped[dd$Location.Description == "TRAILER" | dd$Location.Description == "TRUCK" | dd$Location.Description == "VEHICLE - DELIVERY TRUCK"| dd$Location.Description ==  "VEHICLE - OTHER RIDE SERVICE"| dd$Location.Description ==  "VEHICLE NON-COMMERCIAL" | dd$Location.Description == "VEHICLE-COMMERCIAL" | dd$Location.Description == "CTA BUS" | dd$Location.Description =="DELIVERY TRUCK" | dd$Location.Description =="TAXICAB" | dd$Location.Description == "CTA TRAIN" | dd$Location.Description == "OTHER COMMERCIAL TRANSPORTATION"]<- "Vehicle" 

dd$Location_grouped[dd$Location.Description == "CTA GARAGE / OTHER PROPERTY" | dd$Location.Description =="DRIVEWAY - RESIDENTIAL" | dd$Location.Description =="GARAGE" | dd$Location.Description =="HOUSE" | dd$Location.Description == "PORCH" | dd$Location.Description =="RESIDENCE" | dd$Location.Description == "RESIDENCE PORCH/HALLWAY"| dd$Location.Description == "RESIDENCE-GARAGE"| dd$Location.Description == "RESIDENTIAL YARD (FRONT/BACK)"| dd$Location.Description == "YARD"]<- "Residence"
```

Next, we drop all the variable we do not need.

```{r, echo=TRUE}
# Drop all variables we are not interested in
dd <- dd[, -c(1:8, 10:11, 13:18)]
```

Then, we clean the dataset of missing values.

```{r, echo=TRUE}
# Remove NAs
dd <- dd[complete.cases(dd),]
```

Finally, the data is ready for the explotation.

```{r, echo=TRUE}
head(dd)
```

## Data exploration

### Univariable analysis

#### Crimes evolution 

The number of crimes in Chicago has decrease dramatically per year from 200x until 2015.

```{r fig, fig.cap="Crimes evolution", echo=FALSE, message=FALSE, warning=FALSE}
# Plot the graph 
ggplot(dd, aes(x=Month_Yr))+geom_line(stat="count") + labs(y = "Number of crimes") + theme_minimal() + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
```

Except for the deceptive practice, all the crimes have decresead in more or less grade. 

```{r, include=FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

```{r, fig2, fig.cap="Evolution per type of crime", echo=FALSE, message=FALSE, warning=FALSE}
# Create a graph with each type evolution

p1 <- ggplot(subset(dd, Type_grouped == "Theft"), aes(x=Month_Yr))+geom_line(stat="count")+
labs(y = "Theft") + theme_minimal() + theme(axis.title.x=element_blank()) + theme(text = element_text(size=8))

p2 <- ggplot(subset(dd, Type_grouped == "Batery"), aes(x=Month_Yr))+geom_line(stat="count")+
labs(y = "Batery") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p3 <- ggplot(subset(dd, Type_grouped == "Criminal damage"), aes(x=Month_Yr))+geom_line(stat="count")+
labs(y = "Criminal\n damage") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p4 <- ggplot(subset(dd, Type_grouped == "Narcotics"), aes(x=Month_Yr))+geom_line(stat="count")+
labs(y = "Narcotics") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p5 <- ggplot(subset(dd, Type_grouped == "Others"), aes(x=Month_Yr))+geom_line(stat="count")+
labs(y = "Others") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p6 <- ggplot(subset(dd, Type_grouped == "Other offense"), aes(x=Month_Yr))+geom_line(stat="count")+
labs(y = "Other\n offense") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p7 <- ggplot(subset(dd, Type_grouped == "Assault"),  aes(x=Month_Yr))+geom_line(stat="count") +
labs(y = "Assault") + theme_minimal() + theme(axis.title.x=element_blank()) + theme(text = element_text(size=8))

p8 <- ggplot(subset(dd, Type_grouped == "Burglary"), aes(x=Month_Yr))+geom_line(stat="count")+
labs(y = "Burglary") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p9 <- ggplot(subset(dd, Type_grouped == "Robery"), aes(x=Month_Yr))+geom_line(stat="count")+
labs(y = "Robery") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

p10 <- ggplot(subset(dd, Type_grouped == "Deceptive practice"), aes(x=Month_Yr))+geom_line(stat="count")+
labs(y = "Deceptive\n practice") + theme_minimal() + theme(axis.title.x=element_blank())  + theme(text = element_text(size=8))

multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, cols=2) 
```

#### Crime per Hour

The crimes are concentrated in hours

```{r, fig3, fig.cap="Crimes per hour", echo=FALSE}
# Plot the graph 
ggplot(dd, aes(x=hour))+geom_bar(stat="count")+
  xlab("mon")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Hour", y = "Number of crimes") + theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
```

#### Crime per weekday

Friday concentrated most of the crimes, percentage?

```{r fig4, fig.cap="Crimes per weekday", echo=FALSE}
# Plot the graph 
ggplot(dd, aes(x=weekday))+geom_bar(stat="count")+
  xlab("mon")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Day", y = "Number of crimes") + theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
```

#### Crime per month

Summer is in difference the period with more crimes recorded.

```{r fig5, fig.cap="Crimes per month", echo=FALSE}
# Plot the graph 
ggplot(dd, aes(x=mon))+ geom_bar(stat="count") +
  xlab("mon")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Month", y = "Number of crimes") + theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
```

#### Type of crimes

Per type of crime Theft is in difference the biggest number. Change the scientifyc number.

```{r fig6, fig.cap="Crimes per type", echo=FALSE}
options(scipen=10000)
dd_agg <- aggregate(count ~ Type_grouped, data = dd, FUN = sum)
dd_agg$Type_grouped <- factor(dd_agg$Type_grouped, levels = dd_agg$Type_grouped[order(-dd_agg$count)])
ggplot(dd_agg, aes(x = Type_grouped, y = count)) + theme_minimal() + geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(x = "Type of crimes", y = "Number of crimes")+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
```

#### Location of crimes

These crimes are concentrated in Streets, give percentage.

```{r fig7, fig.cap="Crimes per location", echo=FALSE}
# create aggregated object
dd_agg_l <- aggregate(count ~ Location_grouped, data = dd, FUN = sum)
# order values
dd_agg_l$Location_grouped <- factor(dd_agg_l$Location_grouped, levels = dd_agg_l$Location_grouped[order(-dd_agg_l$count)])
# plot the graph 
ggplot(dd_agg_l, aes(x = Location_grouped, y = count)) + theme_minimal() + geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(x = "Location", y = "Number of crimes")+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
```

#### Crimes per districts

Per districts the most dangerous are 8. 

```{r fig8, fig.cap="Crimes per district", include=FALSE}
# Remove values districts 21 and 31
dd_sub <- subset(dd, District!="21" & District!="31")
# create the aggregated object
dd_agg_d <- aggregate(count ~ District, data = dd_sub, FUN = sum)
# order values
dd_agg_d$District <- factor(dd_agg_d$District, levels = dd_agg_d$District[order(-dd_agg_d$count)])
# Plot the graph 
ggplot(dd_agg_d, aes(x=District, y = count)) + theme_minimal() + geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(x = "District", y = "Number of crimes") + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank())
```

### Multivariable analysis

The multiple analysis focuses on type of crime crossed with hour, location and district.

#### Type of crime vs hour

Are some types of crimes more likely to happen in specific time of the day?

The most dangerous hours per Thefth are 00 and 12.

```{r fig9, fig.cap="Type of crime vs hour", echo=FALSE}
# add variable: performance
count_h <- aggregate(count ~ Type_grouped + hour, data = dd, FUN = sum)

p1 <- ggplot(data = count_h, aes(x = hour, y = Type_grouped)) + geom_tile(aes(fill = count), color = "white") +
  scale_fill_gradient(low = "white", high = "black") 
p1 + theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank()) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 6),
        axis.text.y = element_text(size= 8),
        axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) 
```

#### Type of crime vs location

Are some types of crimes more likely to happen in specific locations?

Street is particularly important for Theft.

```{r fig11, fig.cap="Type of crime vs location", echo=FALSE}
# add variable: performance
count_l <- aggregate(count ~  Type_grouped + Location_grouped, data = dd, FUN = sum)

p2 <- ggplot(data = count_l, aes(x = Location_grouped, y = Type_grouped)) + geom_tile(aes(fill = count), color = "white") +
  scale_fill_gradient(low = "white", high = "black") 
p2+ theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank()) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 6),
        axis.text.y = element_text(size= 8),
        axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) 
```

#### Type of crime vs district

Are some types of crimes more likely to happen in specific distrits?

Narcotics in district 11 is crealy a problem.

```{r fig10, fig.cap="Type of crime vs district", echo=FALSE}
# add variable: performance
count_d <- aggregate(count ~ Type_grouped + District, data = dd_sub, FUN = sum)

p3<-ggplot(data = count_d, aes(x = District, y = Type_grouped)) +
  geom_tile(aes(fill = count), color = "white") +
  scale_fill_gradient(low = "white", high = "black")  
p3+ theme_minimal()+ theme(axis.title.x=element_blank()) + theme(axis.title.y=element_blank()) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 6),
        axis.text.y = element_text(size= 8),
        axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) 
```

# Conclusions
