---
title: 'Crime in Chicago: Data Analysis and Visualizations using R'
author: '201081646'
output:
  word_document:
    fig_caption: yes
  pdf_document: default
  html_document:
    df_print: paged
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
# load libraries
library(ggplot2)
library(ggmap)
library(lubridate)
library(scales)
library(zoo)
```

# Introduction

```{r, message=FALSE, warning=FALSE, include=FALSE}
# load libraries
library(ggplot2)
library(ggmap)
library(lubridate)
library(scales)
library(zoo)
library(dplyr)
```

# Introduction

This is the first assessment for the **Statistical Theory and Methods module**. 

Its objective is to: 

- Summarise a sample of dataset.

- Highlight key finding.

# Data and methods

The dataset we use is a random sample of 500,000 rows of the original data which come from https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2.

```{r eval=TRUE, echo=TRUE}
# read csv in R
dd=read.csv("http://www1.maths.leeds.ac.uk/~charles/math5741/crime.csv",header=T)
```

Explain that all the code is in an appendix.

Also that this report has been done with Rmarkdown.

# Results

## Data preparation

First, we have a look at the variables we got.

```{r, echo=TRUE}
names(dd)
```

We are interested in 8 of them: 
- `Date`,
- `Primary.Type`, 
- `Location.Description`,
- `Arrest`,
- `Domestic`, 
- `District`.

Due to time limitation and lenght constraint. WE focus on some of them and drop the rest of them.

```{r, include=FALSE}
# Drop all variables we are not interested in
dd <- dd[, -c(1:2, 4:5, 7, 11, 13:18)]
```

Secondly, we clean the dataser of missing values - drop the NAs.

```{r, include=FALSE}
# Remove missing values
dd <- dd[complete.cases(dd),]
```

Third, we create new variables:
- `count`,
- `hour`, 
- `Month_Yr`,
- `Month`,
- `weekday`). 

And we give them the right format for explotation.

```{r, include=FALSE}
# Create a variable count with value 1
dd$count <- 1

# Convert Date from factor to date
dd$Date <- mdy_hms(dd$Date)

# Extract hour from Date
dd$hour <- substring(dd$Date, 12,13)

# Drop time from Date
dd$Date <- as.Date(dd$Date, format="%m/%d/%Y")

# Create a month year
dd$Month_Yr <- format(as.Date(dd$Date), "%Y-%m")

# Extract month - the mon component of a POSIXlt object is the numeric month (0-11 starting on January)
dd$mon <- as.POSIXlt(dd$Date)$mon
# Convert weekday from character to factor
dd$mon <- c("January", "February", "March", "April", "May", "June", "July", "Agust", "September", "October", "November", "Desember")[as.POSIXlt(dd$Date)$mon + 1]

# Extract weekdays - the wday component of a POSIXlt object is the numeric weekday (0-6 starting on Sunday)
dd$weekday <- as.POSIXlt(dd$Date)$wday
dd$weekday <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", 
    "Friday", "Saturday")[as.POSIXlt(dd$Date)$wday + 1]
```

```{r, include=FALSE}
# convert all variables in factor
dd$hour <- as.factor(dd$hour)
dd$Month_Yr <- as.Date(as.yearmon(dd$Month_Yr, "%Y-%m"))
dd$mon <- as.factor(dd$mon)
dd$weekday <- as.factor(dd$weekday)
dd$Primary.Type <- as.factor(dd$Primary.Type)
dd$Location.Description <- as.factor(dd$Location.Description)
dd$District<- as.factor(dd$District)
```

We group labels of variables type and location of accidents in bigger categories.

Then we manually groups `Primery.Type` categories. 

```{r, include=FALSE}
# Group Primary.Type categories
dd$Type_grouped[dd$Primary.Type == "THEFT" | dd$Primary.Type == "MOTOR VEHICLE THEFT" ] <- "Theft"

dd$Type_grouped[dd$Primary.Type == "BATTERY"] <- "Batery"

dd$Type_grouped[dd$Primary.Type == "CRIMINAL DAMAGE"] <- "Criminal damage"

dd$Type_grouped[dd$Primary.Type == "NARCOTICS" | dd$Primary.Type == "OTHER NARCOTIC VIOLATION"] <- "Narcotics"

dd$Type_grouped[dd$Primary.Type == "OTHER OFFENSE"] <- "Other offense"

dd$Type_grouped[dd$Primary.Type == "ASSAULT"] <- "Assault"

dd$Type_grouped[dd$Primary.Type == "BURGLARY"] <- "Burglary"

dd$Type_grouped[dd$Primary.Type == "ROBBERY" ] <- "Robery"

dd$Type_grouped[dd$Primary.Type == "DECEPTIVE PRACTICE"] <- "Deceptive practice"

dd$Type_grouped[dd$Primary.Type == "ARSON" | dd$Primary.Type == "CONCEALED CARRY LICENSE VIOLATION" | dd$Primary.Type == "CRIMINAL TRESPASS" | dd$Primary.Type == "GAMBLINGS" | dd$Primary.Type == "HUMAN TRAFFICKING" | dd$Primary.Type == "INTERFERENCE WITH PUBLIC OFFICER" | dd$Primary.Type == "INTIMIDATION" | dd$Type == "KIDNAPPING" | dd$Type == "LIQUOR LAW VIOLATION" | dd$Primary.Type == "NON-CRIMINAL" | dd$Primary.Type == "NON - CRIMINAL" | dd$Primary.Type == "OBSCENITY" | dd$Primary.Type == "OFFENSE INVOLVING CHILDREN"| dd$Primary.Type == "PROSTITUTION"| dd$Primary.Type == "PUBLIC INDECENCY"| dd$Primary.Type == "PUBLIC PEACE VIOLATION"| dd$Primary.Type == "STALKING"| dd$Primary.Type == "WEAPONS VIOLATION"| dd$Primary.Type == "HOMICIDE"| dd$Primary.Type == "CRIM SEXUAL ASSAULT" | dd$Primary.Type == "SEX OFFENSE"] <- "Others"

table(dd$Type_grouped)
```

We do the same with `Location.Description`.

```{r, include=FALSE}
# Group Location.Description categories
dd$Location_grouped[dd$Location.Description == "\"SCHOOL- PRIVATE- BUILDING\"" | dd$Location.Description ==  "\"SCHOOL- PRIVATE- GROUNDS\"" | dd$Location.Description == "\"SCHOOL- PUBLIC- BUILDING\"" | dd$Location.Description == "\"SCHOOL- PUBLIC- GROUNDS\"" | dd$Location.Description == "COLLEGE/UNIVERSITY GROUNDS" | dd$Location.Description == "COLLEGE/UNIVERSITY RESIDENCE HALL"| dd$Location.Description == "SCHOOL YARD"]<- "School or university"

dd$Location_grouped[dd$Location.Description == "" | dd$Location.Description == "ABANDONED BUILDING"| dd$Location.Description == "AIRCRAFT" | dd$Location.Description == "ANIMAL HOSPITAL" | dd$Location.Description == "ATHLETIC CLUB" | dd$Location.Description == "AUTO" | dd$Location.Description == "BASEMENT"  | dd$Location.Description == "BOAT/WATERCRAFT" | dd$Location.Description == "CHA GROUNDS" | dd$Location.Description == "CHA HALLWAY"  | dd$Location.Description == "CHA HALLWAY/STAIRWELL/ELEVATOR" | dd$Location.Description == "CHURCH" | dd$Location.Description == "CHURCH/SYNAGOGUE/PLACE OF WORSHIP"| dd$Location.Description == "COIN OPERATED MACHINE"| dd$Location.Description == "CONSTRUCTION SITE"| dd$Location.Description == "OTHER"| dd$Location.Description == "CHURCH"| dd$Location.Description == "OTHER RAILROAD PROP / TRAIN DEPOT" | dd$Location.Description =="SEWER" | dd$Location.Description =="STAIRWELL"| dd$Location.Description == "VACANT LOT" | dd$Location.Description =="VACANT LOT/LAND"| dd$Location.Description == "VESTIBULE" | dd$Location.Description =="WOODED AREA" | dd$Location.Description == "CTA STATION" | dd$Location.Description == "CTA BUS STOP"| dd$Location.Description == "CTA TRACKS - RIGHT OF WAY"  | dd$Location.Description == "AIRPORT BUILDING NON-TERMINAL - NON-SECURE AREA" | dd$Location.Description == "AIRPORT BUILDING NON-TERMINAL - SECURE AREA" | dd$Location.Description == "AIRPORT EXTERIOR - NON-SECURE AREA" | dd$Location.Description == "AIRPORT EXTERIOR - SECURE AREA" | dd$Location.Description == "AIRPORT PARKING LOT" | dd$Location.Description ==  "AIRPORT TERMINAL LOWER LEVEL - NON-SECURE AREA" | dd$Location.Description == "AIRPORT TERMINAL LOWER LEVEL - SECURE AREA" | dd$Location.Description == "AIRPORT TERMINAL MEZZANINE - NON-SECURE AREA"| dd$Location.Description == "AIRPORT TERMINAL UPPER LEVEL - NON-SECURE AREA"| dd$Location.Description == "AIRPORT TERMINAL UPPER LEVEL - SECURE AREA"| dd$Location.Description == "AIRPORT TRANSPORTATION SYSTEM (ATS)"| dd$Location.Description == "AIRPORT VENDING ESTABLISHMENT" | dd$Location.Description == "AIRPORT/AIRCRAFT" |dd$Location.Description == "DAY CARE CENTER"| dd$Location.Description == "COMMERCIAL / BUSINESS OFFICE"| dd$Location.Description == "FACTORY" | dd$Location.Description =="FACTORY/MANUFACTURING BUILDING" | dd$Location.Description =="FEDERAL BUILDING"| dd$Location.Description == "FIRE STATION"| dd$Location.Description == "GOVERNMENT BUILDING/PROPERTY"| dd$Location.Description == "HOSPITAL"| dd$Location.Description == "HOSPITAL BUILDING/GROUNDS"| dd$Location.Description == "JAIL / LOCK-UP FACILITY"| dd$Location.Description == "LIBRARY"| dd$Location.Description == "MOVIE HOUSE/THEATER"  | dd$Location.Description =="NURSING HOME/RETIREMENT HOME"| dd$Location.Description == "POOL ROOM" | dd$Location.Description =="SPORTS ARENA/STADIUM"| dd$Location.Description == "WAREHOUSE" | dd$Location.Description == "BANK"| dd$Location.Description == "CREDIT UNION"| dd$Location.Description == "CURRENCY EXCHANGE"| dd$Location.Description == "SAVINGS AND LOAN"]<- "Others"

dd$Location_grouped[dd$Location.Description == "ALLEY" | dd$Location.Description == "BOWLING ALLEY" | dd$Location.Description == "CHA BREEZEWAY" | dd$Location.Description =="HALLWAY"]<- "Alley" 

dd$Location_grouped[dd$Location.Description == "APARTMENT"| dd$Location.Description == "CHA APARTMENT"]<- "Apartment"

dd$Location_grouped[dd$Location.Description == "APPLIANCE STORE" | dd$Location.Description == "BARBERSHOP" | dd$Location.Description == "CAR WASH" | dd$Location.Description == "CLEANING STORE" | dd$Location.Description ==  "CONVENIENCE STORE" | dd$Location.Description =="DEPARTMENT STORE" | dd$Location.Description =="DRUG STORE"| dd$Location.Description == "GARAGE/AUTO REPAIR"| dd$Location.Description == "GAS STATION"| dd$Location.Description == "GAS STATION DRIVE/PROP." | dd$Location.Description =="GROCERY FOOD STORE"| dd$Location.Description == "MEDICAL/DENTAL OFFICE" | dd$Location.Description =="NEWSSTAND"| dd$Location.Description == "OFFICE"| dd$Location.Description == "PAWN SHOP" | dd$Location.Description =="RETAIL STORE"| dd$Location.Description == "SMALL RETAIL STORE"]<- "Store/small business"

dd$Location_grouped[dd$Location.Description == "ATM (AUTOMATIC TELLER MACHINE)" | dd$Location.Description == "BRIDGE"| dd$Location.Description == "DRIVEWAY"| dd$Location.Description == "GANGWAY"| dd$Location.Description == "HIGHWAY/EXPRESSWAY"| dd$Location.Description == "LAKEFRONT/WATERFRONT/RIVERBANK"| dd$Location.Description == "SIDEWALK" | dd$Location.Description == "STREET"]<- "Street"

dd$Location_grouped[dd$Location.Description == "BAR OR TAVERN"| dd$Location.Description == "HOTEL"| dd$Location.Description == "HOTEL/MOTEL"| dd$Location.Description == "RESTAURANT"| dd$Location.Description == "TAVERN"| dd$Location.Description == "TAVERN/LIQUOR STORE"]<- "Restaurant/bar/hotel"

dd$Location_grouped[dd$Location.Description == "CHA PARKING LOT" | dd$Location.Description =="PARK PROPERTY" | dd$Location.Description =="PARKING LOT"| dd$Location.Description == "PARKING LOT/GARAGE(NON.RESID.)" | dd$Location.Description =="POLICE FACILITY/VEH PARKING LOT"]<- "Parking lot"

dd$Location_grouped[dd$Location.Description == "TRAILER" | dd$Location.Description == "TRUCK" | dd$Location.Description == "VEHICLE - DELIVERY TRUCK"| dd$Location.Description ==  "VEHICLE - OTHER RIDE SERVICE"| dd$Location.Description ==  "VEHICLE NON-COMMERCIAL" | dd$Location.Description == "VEHICLE-COMMERCIAL" | dd$Location.Description == "CTA BUS" | dd$Location.Description =="DELIVERY TRUCK" | dd$Location.Description =="TAXICAB" | dd$Location.Description == "CTA TRAIN" | dd$Location.Description == "OTHER COMMERCIAL TRANSPORTATION"]<- "Vehicle" 

dd$Location_grouped[dd$Location.Description == "CTA GARAGE / OTHER PROPERTY" | dd$Location.Description =="DRIVEWAY - RESIDENTIAL" | dd$Location.Description =="GARAGE" | dd$Location.Description =="HOUSE" | dd$Location.Description == "PORCH" | dd$Location.Description =="RESIDENCE" | dd$Location.Description == "RESIDENCE PORCH/HALLWAY"| dd$Location.Description == "RESIDENCE-GARAGE"| dd$Location.Description == "RESIDENTIAL YARD (FRONT/BACK)"| dd$Location.Description == "YARD"]<- "Residence"

table(dd$Location_grouped)
```

Finally, we show the data prepared for explotation.

```{r, echo=TRUE}
head(dd[dd$VAR1==4,],6)
```

## Data exploration

This section explores the data visually to identify the key patterns.

### Univariable analysis

#### Evolution crime 

Apint each of the lines in different colour.

```{r, include=FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

It´s too similar to kaggle. Try using arrents. Maybe not analysing the evlution per each.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Plot the graph 
ggplot(dd, aes(x=Month_Yr))+geom_line(stat="count") + labs(y = "Number of crimes") + theme(axis.title.x=element_blank())

# Create a graph with each type evolution
p1 <- ggplot(subset(dd, Type_grouped == "Assault"), aes(x=Month_Yr))+geom_line(stat="count", colour="red")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

p2 <- ggplot(subset(dd, Type_grouped == "Batery"), aes(x=Month_Yr))+geom_line(stat="count", colour="#000199")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

p3 <- ggplot(subset(dd, Type_grouped == "Burglary"), aes(x=Month_Yr))+geom_line(stat="count", colour="yellow")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

p4 <- ggplot(subset(dd, Type_grouped == "Criminal damage"), aes(x=Month_Yr))+geom_line(stat="count", colour="green")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

p5 <- ggplot(subset(dd, Type_grouped == "Deceptive practice"), aes(x=Month_Yr))+geom_line(stat="count", colour="grey")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

p6 <- ggplot(subset(dd, Type_grouped == "Narcotics"), aes(x=Month_Yr))+geom_line(stat="count", colour="pink")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

p7 <- ggplot(subset(dd, Type_grouped == "Other offense"), aes(x=Month_Yr))+geom_line(stat="count", colour="blue")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

p8 <- ggplot(subset(dd, Type_grouped == "Others"), aes(x=Month_Yr))+geom_line(stat="count", colour="magenta")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

p9 <- ggplot(subset(dd, Type_grouped == "Robery"), aes(x=Month_Yr))+geom_line(stat="count", colour="orange")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

p10 <- ggplot(subset(dd, Type_grouped == "Theft"), aes(x=Month_Yr))+geom_line(stat="count", colour="brown")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, cols=2)
```

Order all the maps from more to less.

#### Records by time

##### Hour

```{r, echo=FALSE}
# Plot the graph 
ggplot(dd, aes(x=hour))+geom_bar(stat="count")+
  xlab("mon")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Hour", y = "Number of crimes") + coord_flip()
```

##### Weekday - Order days

```{r, echo=FALSE}
# Plot the graph 
ggplot(dd, aes(x=weekday))+geom_bar(stat="count")+
  xlab("mon")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Day", y = "Number of crimes") + coord_flip()
```

##### Month - Order months

```{r, echo=FALSE}
# Plot the graph 
ggplot(dd, aes(x=mon))+ geom_bar(stat="count") +
  xlab("mon")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Month", y = "Number of crimes") + coord_flip()
```

#### Type of crimes

```{r, echo=FALSE}
# Plot the graph 
ggplot(dd, aes(x=Type_grouped))+ geom_bar(stat="count") +
  xlab("mon")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Month", y = "Number of crimes") + coord_flip()
```

#### Location of crimes

```{r, echo=FALSE}
# Plot the graph 
ggplot(dd, aes(x=Location_grouped))+ geom_bar(stat="count") +
  xlab("mon")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Month", y = "Number of crimes") + coord_flip()
```

#### Crimes per districts (map?)

```{r, include=FALSE}
# Remove values districts 21 and 31
dd_sub <- subset(dd, District!="21" & District!="31")
```

```{r, echo=FALSE}
# Plot the graph 
ggplot(dd_sub, aes(x=District))+ geom_bar(stat="count") +
  xlab("mon")+ theme(axis.text.x = element_text(angle = 0, hjust = 1)) + labs(x = "Month", y = "Number of crimes") + coord_flip()
```

### Two variables analysis

#### Hour vs Type of crime

```{r, echo=FALSE}
# add variable: performance
count_h <- aggregate(count ~ Type_grouped + hour, data = dd, FUN = sum)

p1 <- ggplot(data = count_h, aes(x = hour, y = Type_grouped)) + geom_tile(aes(fill = count), color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  ylab("Crime") +
  xlab("Location") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(fill = "Number of crimes")
p1
```

#### District vs Type of crimes

```{r, echo=FALSE}
# add variable: performance
count_d <- aggregate(count ~ Type_grouped + District, data = dd_sub, FUN = sum)

p2<-ggplot(data = count_d, aes(x = District, y = Type_grouped)) +
  geom_tile(aes(fill = count), color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  ylab("Crime") +
  xlab("District") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(fill = "Number of crimes")
p2
```

#### Location vs Type of crimes

```{r, echo=FALSE}
# add variable: performance
count_l <- aggregate(count ~ Type_grouped + Location_grouped, data = dd, FUN = sum)

p3 <- ggplot(data = count_l, aes(x = Location_grouped, y = Type_grouped)) + geom_tile(aes(fill = count), color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  ylab("Crime") +
  xlab("Location") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(fill = "Number of crimes")
p3
```

# Discussion

# Conclusions

# Code Appendix

```{r ref.label=knitr::all_labels(), echo = T, eval = F}
```